\pdfbookmark[1]{Abstract}{Abstract} % Bookmark name visible in a PDF viewer

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract} % Abstract name
How do circuits in the mammalian cerebral cortex encode properties of
the sensory environment in a way that can drive adaptive behavior?
This question is fundamental to neuroscience, but it has been very
difficult to approach directly. Various computational and theoretical
models can explain a wide range of phenomena observed in the primary
visual cortex (V1), including the anatomical organization of its
circuits, the development of functional properties like orientation
tuning, and behavioral effects like surround modulation. However, so
far no model has been able to bridge these levels of description to
explain how the machinery that develops directly affects
behavior. Bridging these levels is important, because phenomena at any
one specific level can have many possible explanations, but there are
far fewer possibilities to consider once all of the available evidence
is taken into account.

In this thesis we integrate the information gleaned about cortical
development, circuit and cell-type specific interactions, and
anatomical, behavioral and electrophysiological measurements, to
develop a computational model of V1 that is constrained enough to make
predictions across multiple levels of description. Through a series of
models incorporating increasing levels of biophysical detail and
becoming increasingly better constrained, we are able to make detailed
predictions for the types of mechanistic interactions required for
robust development of cortical maps that have a realistic anatomical
organization, and thereby gain insight into the computations performed
by the primary visual cortex.

The initial models focus on how existing anatomical and
electrophysiological knowledge can be integrated into previously
abstract models to give a well-grounded and highly constrained account
of the emergence of pattern-specific tuning in the primary visual
cortex. More detailed models then address the interactions between
specific excitatory and inhibitory cell classes in V1, and what role
each cell type may play during development and function. Finally, we
demonstrate how these cell classes come together to form a circuit
that gives rise not only to robust development but also the
development of realistic lateral connectivity patterns.  Crucially,
these patterns reflect the statistics of the visual environment to
which the model was exposed during development. This property allows
us to explore how the model is able to capture higher-order
information about the environment and use that information to optimize
neural coding and aid the processing of complex visual tasks.

Using this model we can make a number of very specific predictions
about the mechanistic workings of the brain. Specifically, the model
predicts a crucial role of parvalbumin-expressing interneurons in
robust development and divisive normalization, while it implicates
somatostatin immunoreactive neurons in mediating longer range and
feature-selective suppression. The model also makes predictions about
the role of these cell classes in efficient neural coding and under
what conditions the model fails to organize. In particular, we show
that a tight coupling of activity between the principal excitatory
population and the parvalbumin population is central to robust and
stable responses and organization, which may have implications for a
variety of diseases where parvalbumin interneuron function is
impaired, such as schizophrenia and autism. Further the model explains
the switch from facilitatory to suppressive surround modulation
effects as a simple by-product of the facilitating response function
of long-range excitatory connections targeting a specialized class of
inhibitory interneurons. Finally, the model allows us to make
predictions about the statistics that are encoded in the extensive
network of long-range intra-areal connectivity in V1, suggesting that
even V1 can capture high-level statistical dependencies in the visual
environment.

The final model represents a comprehensive and well constrained model
of the primary visual cortex, which for the first time can relate the
physiological properties of individual cell classes to their role in
development, learning and function. While the model is specifically
tuned for V1, all mechanisms introduced are completely general, and
can be used as a general cortical model, useful for studying phenomena
across the visual cortex even the cortex as a whole. This work is also
highly relevant for clinical neuroscience, as the cell types studied
here have been implicated in neurological disorders as wide ranging as
autism, schizophrenia and Parkinson's disease.
\endgroup
